### 2018-07-06 Operads_rewrite.tex
Partie 1 chantier (suite)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The magmatic operad, quotients, and rewrite relations}
\label{sec:operad_Mag}
We set in this preliminary section our notations about operads. We also
provide a definition for the magmatic operad and introduce tools to
handle with some of its quotients involving rewrite systems on syntax
trees.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonsymmetric operads}
A \Def{nonsymmetric operad in the category of sets}, or a
\Def{nonsymmetric operad} for short, is a graded set
\begin{equation}
    \Oca = \bigsqcup_{n \geq 1} \Oca(n)
\end{equation}
together with maps
\begin{equation}
    \circ_i : \Oca(n) \times \Oca(m) \to \Oca(n + m - 1),
    \qquad 1 \leq i \leq n, 1 \leq m,
\end{equation}
called \Def{partial compositions}, and a distinguished element
$\Unit \in \Oca(1)$, the \Def{unit} of $\Oca$. This data has to satisfy,
for any $x, y, z \in \Oca$, the three relations
\begin{subequations}
\begin{equation} \label{equ:operad_axiom_1}
    (x \circ_i y) \circ_{i + j - 1} z = x \circ_i (y \circ_j z),
    \qquad
    1 \leq i \leq |x|, 1 \leq j \leq |y|,
\end{equation}
\begin{equation} \label{equ:operad_axiom_2}
    (x \circ_i y) \circ_{j + |y| - 1} z = (x \circ_j z) \circ_i y,
    \qquad
    1 \leq i < j \leq |x|,
\end{equation}
\begin{equation} \label{equ:operad_axiom_3}
    \Unit \circ_1 x = x = x \circ_i \Unit,
    \qquad 1 \leq i \leq |x|.
\end{equation}
\end{subequations}
Since we consider in this work only nonsymmetric operads, we shall call
these simply \Def{operads}.
\medbreak

Let us provide some elementary definitions and notations about operads.
If $x$ is an element of $\Oca$ such that $x \in \Oca(n)$ for an
$n \geq 1$, the \Def{arity} $|x|$ of $x$ is $n$. The
\Def{complete composition maps} of $\Oca$ are the map
\begin{equation}
    \circ : \Oca(n) \times \Oca(m_1) \times \dots \times \Oca(m_n)
    \to \Oca(m_1 + \dots + m_n)
\end{equation}
defined, for any $x \in \Oca(n)$ and $y_1, \dots, y_n \in \Oca$, by
\begin{equation}
    x \circ [y_1, \dots, y_n] :=
    (\dots ((x \circ_n y_n) \circ_{n - 1} y_{n - 1}) \dots) \circ_1 y_1.
\end{equation}
If $\Oca_1$ and $\Oca_2$ are two operads, a map
$\phi : \Oca_1 \to \Oca_2$ is an \Def{operad morphism} is it respects
the arities, sends the unit of $\Oca_1$ to the unit of $\Oca_2$, and
commutes with partial composition maps. We say that $\Oca_2$ is a
\Def{suboperad} of $\Oca_1$ if $\Oca_2$ is a subset of $\Oca_1$
containing the unit of $\Oca_1$ and the partial composition maps of
$\Oca_2$ are the ones of $\Oca_1$ restricted on $\Oca_2$. For any subset
$\GeneratingSet$ of $\Oca$, the \Def{operad generated} by
$\GeneratingSet$ is the smallest suboperad $\Oca^\GeneratingSet$ of
$\Oca$ containing $\GeneratingSet$. When $\Oca = \Oca^\GeneratingSet$,
we say that $\GeneratingSet$ is a \Def{generating set} of $\Oca$. In
this case, and when additionally $\GeneratingSet$ is minimal among all
the subsets of $\Oca$ satisfying this property, $\GeneratingSet$ is a
\Def{minimal} generating set of $\Oca$. An \Def{operad congruence} is an
equivalence relation $\Congr$ on $\Oca$ such that $\Congr$ respects the
arities, and for any $x, x', y, y' \in \Oca$ such that $x \Congr x'$ and
$y \Congr y'$, $x \circ_i y$ is $\Congr$-equivalent to $x' \circ_i y'$
for any valid integer $i$. Given an operad congruence $\Congr$, the
\Def{quotient operad} $\Oca/_{\Congr}$ of $\Oca$ by $\Congr$ is the
operad of all $\Congr$-equivalence classes endowed with the partial
composition maps defined in the obvious way. In the case where all the
sets $\Oca(n)$, $n \geq 1$, are finite, the \Def{Hilbert series}
$\HilbertSeries_\Oca(t)$ of $\Oca$ is the series defined by
\begin{equation}
    \HilbertSeries_\Oca(t) := \sum_{n \geq 1} \# \Oca(n) t^n.
\end{equation}
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Binary trees and magmatic operad}
A \Def{binary tree} is either the \Def{leaf} $\Leaf$ or a pair
$(\Tfr_1, \Tfr_2)$ of binary trees. We use the standard terminology
about binary trees (such as \Def{root}, \Def{internal node},
\Def{left child}, \Def{right child}, {\em etc.}) in this work. Let us
recall the main notions. The \Def{arity} $|\Tfr|$ (resp. \Def{degree}
$\Deg(\Tfr)$) of a binary tree $\Tfr$ is its number of leaves (resp.
internal nodes). We shall draw binary trees the root to the top. For
instance,
\begin{equation}
    \begin{tikzpicture}[xscale=.15,yscale=.18,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-6.75){};
        \node(4)at(4.00,-6.75){};
        \node(6)at(6.00,-4.50){};
        \node(8)at(8.00,-4.50){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-4.50){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-2.25){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(5);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(5.00,2){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
\end{equation}
is the graphical representation of the binary tree
\begin{math}
    ((\Leaf, (\Leaf, \Leaf)), (\Leaf, \Leaf)).
\end{math}
\medbreak

The \Def{magmatic operad} $\Mag$ is the graded set of all binary trees
where $\Mag(n)$, $n \geq 1$, is the set of all binary trees of arity
$n$. The partial composition maps of $\Mag$ are grafting of trees: given
two binary trees $\Tfr$ and $\Sfr$, $\Tfr \circ_i \Sfr$ is the binary
tree obtained by grafting the root of $\Sfr$ onto the $i$th leaf
(numbered from left to right) of $\Tfr$. For instance,
\begin{equation} \label{equ:example_composition_mag}
    \begin{tikzpicture}[xscale=.15,yscale=.18,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-6.75){};
        \node(4)at(4.00,-6.75){};
        \node(6)at(6.00,-4.50){};
        \node(8)at(8.00,-4.50){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-4.50){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-2.25){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(5);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(5.00,2){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \enspace \circ_4 \enspace
    \begin{tikzpicture}[xscale=.17,yscale=.18,Centering]
        \node(0)at(0.00,-4.67){};
        \node(2)at(2.00,-4.67){};
        \node(4)at(4.00,-4.67){};
        \node(6)at(6.00,-4.67){};
        \node[NodeST](1)at(1.00,-2.33){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-2.33){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(3.00,2){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}
    \enspace = \enspace
    \begin{tikzpicture}[xscale=.15,yscale=.14,Centering]
        \node(0)at(0.00,-6.00){};
        \node(10)at(10.00,-12.00){};
        \node(12)at(12.00,-12.00){};
        \node(14)at(14.00,-6.00){};
        \node(2)at(2.00,-9.00){};
        \node(4)at(4.00,-9.00){};
        \node(6)at(6.00,-12.00){};
        \node(8)at(8.00,-12.00){};
        \node[NodeST](1)at(1.00,-3.00){\begin{math}\Product\end{math}};
        \node[NodeST](11)at(11.00,-9.00){\begin{math}\Product\end{math}};
        \node[NodeST](13)at(13.00,-3.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-6.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-9.00){\begin{math}\Product\end{math}};
        \node[NodeST](9)at(9.00,-6.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(5);
        \draw[Edge](10)--(11);
        \draw[Edge](11)--(9);
        \draw[Edge](12)--(11);
        \draw[Edge](13)--(5);
        \draw[Edge](14)--(13);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(9);
        \draw[Edge](8)--(7);
        \draw[Edge](9)--(13);
        \node(r)at(5.00,2.5){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
\end{equation}
is a partial composition in $\Mag$. This leads, by definition, to the
following complete composition maps of $\Mag$. Given $\Tfr \in \Mag(n)$
and $\Sfr_1, \dots, \Sfr_n \in \Mag$,
\begin{math}
    \Tfr \circ [\Sfr_1, \dots, \Sfr_n]
\end{math}
is the binary tree obtained by grafting simultaneously the roots of all
$\Sfr_i$ of the $i$th leaves of $\Tfr$. The number of binary of arity
$n \geq 1$ is the $n$th Catalan number $\Catalan(n)$ and hence, the
Hilbert series of $\Mag$ is
\begin{equation}
    \HilbertSeries_{\Mag}(t)
    = \sum_{n \geq 1} \Catalan(n) t^n
    = \sum_{n \geq 1} \binom{2n - 1}{n - 1} \frac{1}{n} t^n.
\end{equation}
\medbreak

The operad $\Mag$ is the free operad generated by one binary element. It
satisfies the following universality property. Let
\begin{math}
    \GeneratingSet := \GeneratingSet(2) := \{\Product\}
\end{math}
be the graded set containing exactly one element $\Product$ of arity
$2$. For any operad $\Oca$ and any map
\begin{math}
    f : \GeneratingSet(2) \to \Oca(2),
\end{math}
there exists a unique operad morphism $\phi : \Mag \to \Oca$ such that
$f = \phi \circ \Corolla$, where $\Corolla$ is the map sending
$\Product$ to the unique binary tree of degree $1$ (and then, arity
$2$). In other terms, the diagram
\begin{equation}
    \begin{tikzpicture}[xscale=1.3,yscale=1.1,Centering]
        \node(G)at(0,0){\begin{math}\GeneratingSet\end{math}};
        \node(O)at(2,0){\begin{math}\Oca\end{math}};
        \node(AG)at(0,-2){\begin{math}\Mag\end{math}};
        \draw[Map](G)--(O)node[midway,above]{\begin{math}f\end{math}};
        \draw[Injection](G)--(AG)node[midway,left]
            {\begin{math}\Corolla\end{math}};
        \draw[Map,dashed](AG)--(O)node[midway,right]
            {\begin{math}\phi\end{math}};
    \end{tikzpicture}
\end{equation}
commutes.
\medbreak

We now provide some useful tools about binary trees.
Given a binary tree $\Tfr$, we denote by $ \PrefixWord(\Tfr)$ the
\Def{prefix word} of $\Tfr$, that is the word on $\{\Zero, \Two\}$
obtained by a left to right depth-first traversal of $\Tfr$ and by
writing $\Zero$ (resp. $\Two$) when a leaf (resp. an internal node) is
encountered. For instance,
\begin{equation}
    \begin{tikzpicture}[xscale=.15,yscale=.14,Centering]
        \node(0)at(0.00,-6.00){};
        \node(10)at(10.00,-12.00){};
        \node(12)at(12.00,-12.00){};
        \node(14)at(14.00,-6.00){};
        \node(2)at(2.00,-9.00){};
        \node(4)at(4.00,-9.00){};
        \node(6)at(6.00,-12.00){};
        \node(8)at(8.00,-12.00){};
        \node[NodeST](1)at(1.00,-3.00){\begin{math}\Product\end{math}};
        \node[NodeST](11)at(11.00,-9.00){\begin{math}\Product\end{math}};
        \node[NodeST](13)at(13.00,-3.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-6.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-9.00){\begin{math}\Product\end{math}};
        \node[NodeST](9)at(9.00,-6.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(5);
        \draw[Edge](10)--(11);
        \draw[Edge](11)--(9);
        \draw[Edge](12)--(11);
        \draw[Edge](13)--(5);
        \draw[Edge](14)--(13);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(9);
        \draw[Edge](8)--(7);
        \draw[Edge](9)--(13);
        \node(r)at(5.00,2.5){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \xmapsto{\; \PrefixWord \;}
    \Two \Two \Zero \Two \Zero \Zero \Two \Two \Two \Zero \Zero \Two
    \Zero \Zero \Zero.
\end{equation}
The set of all words on $\{\Zero, \Two\}$ is endowed with
the lexicographic order $\leq$ induced by $\Zero < \Two$.
The \Def{left rank} of $\Tfr$ as the number $\LBranch(\Tfr)$ of internal
nodes in the left branch beginning at the root of $\Tfr$. For instance,
\begin{equation}
    \begin{tikzpicture}[xscale=.15,yscale=.14,Centering]
        \node(0)at(0.00,-9.00){};
        \node(10)at(10.00,-9.00){};
        \node(12)at(12.00,-9.00){};
        \node(14)at(14.00,-9.00){};
        \node(2)at(2.00,-12.00){};
        \node(4)at(4.00,-12.00){};
        \node(6)at(6.00,-6.00){};
        \node(8)at(8.00,-9.00){};
        \node[NodeST](1)at(1.00,-6.00){\begin{math}\Product\end{math}};
        \node[NodeST](11)at(11.00,-3.00){\begin{math}\Product\end{math}};
        \node[NodeST](13)at(13.00,-6.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-9.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-3.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](9)at(9.00,-6.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(5);
        \draw[Edge](10)--(9);
        \draw[Edge](11)--(7);
        \draw[Edge](12)--(13);
        \draw[Edge](13)--(11);
        \draw[Edge](14)--(13);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \draw[Edge](5)--(7);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(9);
        \draw[Edge](9)--(11);
        \node(r)at(7.00,2.25){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \xmapsto{\; \LBranch\; } 3.
\end{equation}
Equivalently, $\LBranch(\Tfr)$ is the length of the prefix of
$\PrefixWord(\Tfr)$ containing only the letter $\Two$. A binary tree
$\Sfr$ is a \Def{subtree} of $\Tfr$ if it possible to stack $\Sfr$ onto
$\Tfr$ by possibly superimposing leaves of $\Sfr$ onto internal nodes of
$\Tfr$. More formally, by using the operad $\Mag$, this is equivalent to
the fact that $\Tfr$ expresses as
\begin{equation}
    \Tfr = \Rfr \circ_i
    \left(\Sfr \circ \left[\Rfr_1, \dots, \Rfr_n\right]\right)
\end{equation}
where $\Rfr$ and $\Rfr_1$, \dots, $\Rfr_n$ are binary trees,
$i \in [|\Rfr|]$, and $n$ is the arity of $\Sfr$. When, on the contrary,
$\Sfr$ is not a subtree of $\Tfr$, we say that $\Tfr$ \Def{avoids}
$\Sfr$.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rewrite systems and presentations}
We present here notions about rewrite systems on binary trees. General
notations and notions appear in~\cite{BN98}.
\medbreak

A \Def{rewrite rule} is an ordered pair $(\Sfr, \Sfr')$ of binary trees
such that $|\Sfr| = |\Sfr'|$. A set $S$ of rewrite rules is a binary
relation on $\Mag$ and it shall be denoted by $\Rew$. We denote by
$\Sfr \Rew \Sfr'$ the fact that $(\Sfr, \Sfr') \in \Rew$. In the sequel,
to define a set of rewrite rules $\Rew$, we shall simply list all
the pairs $\Sfr \Rew \Sfr'$ contained in $\Rew$.
\medbreak

If $\Rew$ is a set of rewrite rules, we denote by $\RewContext$ the
\Def{rewrite relation induced} by $\Rew$. Formally we have
\begin{equation}
    \Tfr \circ_i
    \left(\Sfr \circ \left[\Rfr_1, \dots, \Rfr_{n}\right]\right)
    \RewContext
    \Tfr \circ_i
    \left(\Sfr' \circ \left[\Rfr_1, \dots, \Rfr_{n}\right]\right),
\end{equation}
if $\Sfr \Rew \Sfr'$ where $n = |\Sfr|$, and $\Tfr$, $\Rfr_1$, \dots,
$\Rfr_n$ are binary trees. In other words, one has
$\Tfr \RewContext \Tfr'$ if it is possible to obtain $\Tfr'$ from $\Tfr$
by replacing a subtree $\Sfr$ of $\Tfr$ by $\Sfr'$ whenever
$\Sfr \Rew \Sfr'$. For instance, if $\Rew$ is the set of rewrite rules
satisfying
\begin{equation}
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.50){};
        \node(2)at(2.00,-5.25){};
        \node(4)at(4.00,-5.25){};
        \node(6)at(6.00,-1.75){};
        \node[NodeST](1)at(1.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.50){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge,ultra thick](1)--(5);
        \draw[Edge](2)--(3);
        \draw[Edge,ultra thick](3)--(1);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(5.00,1.31){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \enspace \Rew \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-1.75){};
        \node(2)at(2.00,-3.50){};
        \node(4)at(4.00,-5.25){};
        \node(6)at(6.00,-5.25){};
        \node[NodeST](1)at(1.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-3.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge,ultra thick](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge,ultra thick](5)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(1.00,1.31){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}
\end{equation}
one has
\begin{equation}
    \begin{tikzpicture}[xscale=.19,yscale=.14,Centering]
        \node(0)at(0.00,-9.00){};
        \node(10)at(10.00,-12.00){};
        \node(12)at(12.00,-18.00){};
        \node(14)at(14.00,-18.00){};
        \node(16)at(16.00,-15.00){};
        \node(18)at(18.00,-9.00){};
        \node(2)at(2.00,-9.00){};
        \node(20)at(20.00,-9.00){};
        \node(4)at(4.00,-6.00){};
        \node(6)at(6.00,-12.00){};
        \node(8)at(8.00,-12.00){};
        \node[NodeST](1)at(1.00,-6.00){\begin{math}\Product\end{math}};
        \node[NodeST](11)at(11.00,-9.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](13)at(13.00,-15.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](15)at(15.00,-12.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](17)at(17.00,-3.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](19)at(19.00,-6.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-9.00){\begin{math}\Product\end{math}};
        \node[NodeST](9)at(9.00,-6.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](10)--(11);
        \draw[Edge,ultra thick](11)--(9);
        \draw[Edge](12)--(13);
        \draw[Edge](13)--(15);
        \draw[Edge](14)--(13);
        \draw[Edge](15)--(11);
        \draw[Edge](16)--(15);
        \draw[Edge](17)--(5);
        \draw[Edge](18)--(19);
        \draw[Edge](19)--(17);
        \draw[Edge](2)--(1);
        \draw[Edge](20)--(19);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(9);
        \draw[Edge](8)--(7);
        \draw[Edge,ultra thick](9)--(17);
        \node(r)at(5.00,2.25){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \enspace \RewContext \enspace
    \begin{tikzpicture}[xscale=.19,yscale=.14,Centering]
        \node(0)at(0.00,-9.00){};
        \node(10)at(10.00,-9.00){};
        \node(12)at(12.00,-18.00){};
        \node(14)at(14.00,-18.00){};
        \node(16)at(16.00,-15.00){};
        \node(18)at(18.00,-15.00){};
        \node(2)at(2.00,-9.00){};
        \node(20)at(20.00,-15.00){};
        \node(4)at(4.00,-6.00){};
        \node(6)at(6.00,-9.00){};
        \node(8)at(8.00,-9.00){};
        \node[NodeST](1)at(1.00,-6.00){\begin{math}\Product\end{math}};
        \node[NodeST](11)at(11.00,-6.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](13)at(13.00,-15.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](15)at(15.00,-12.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](17)at(17.00,-9.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](19)at(19.00,-12.00)
            {\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-6.00){\begin{math}\Product\end{math}};
        \node[NodeST](9)at(9.00,-3.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](10)--(11);
        \draw[Edge,ultra thick](11)--(9);
        \draw[Edge](12)--(13);
        \draw[Edge](13)--(15);
        \draw[Edge](14)--(13);
        \draw[Edge](15)--(17);
        \draw[Edge](16)--(15);
        \draw[Edge,ultra thick](17)--(11);
        \draw[Edge](18)--(19);
        \draw[Edge](19)--(17);
        \draw[Edge](2)--(1);
        \draw[Edge](20)--(19);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(9);
        \draw[Edge](8)--(7);
        \draw[Edge](9)--(5);
        \node(r)at(5.00,2.25){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}\,.
\end{equation}
\medbreak

Let $\Rew$ be a set of rewrite rules and $\RewContext$ be rewrite
relation induced by $\Rew$. Since $\RewContext$ is in particular a
binary relation on $\Mag$, the classical notations about closures apply
here: we denote by $\RewContextT$ (resp. $\RewContextRT$,
$\RewContextRST$) the transitive (resp. reflexive and transitive,
reflexive, symmetric, and transitive) closure of $\RewContext$.
\medbreak

When $\Tfr_0$, $\Tfr_1$, \dots, $\Tfr_k$ are binary trees such that
\begin{equation}
    \Tfr_0 \RewContext \Tfr_1 \RewContext \cdots \RewContext \Tfr_k,
\end{equation}
we say that $\Tfr_0$ is \Def{rewritable} by $\RewContext$ into $\Tfr_k$
in $k$ \Def{steps}. When there is no infinite chain
\begin{equation} \label{equ:infinite_chain}
    \Tfr_0 \RewContext \Tfr_2 \RewContext \Tfr_3 \RewContext \Tfr_4
    \cdots
\end{equation}
we say that $\RewContext$ is terminating. To establish the termination
of a rewrite relation is terminating, we will use the following
criterion.
\medbreak

\begin{Lemma}\label{lem:prefix_word_termination}
    Let $\Rew$ be a set of rewrite rules on $\Mag$. If for any
    $\Tfr, \Tfr' \in \Mag$ such that $\Tfr \Rew \Tfr'$ one has
    $\PrefixWord(\Tfr) > \PrefixWord(\Tfr')$, then the rewrite relation
    induced by $\Rew$ is terminating.
\end{Lemma}
\begin{proof}
    \Todo{à faire}
    % 1. Montrer la compatibilité des compo aux mots prefixes
    % 2. Supposer qu'il existe une chaine infinie de reecritures
    % 3. Absurde car il n'existe qu'un nombre fini de mots sur {0, 2}.
\end{proof}
\medbreak

A \Def{normal form} for $\RewContext$ is a binary tree $\Tfr$ such
that for all $\Tfr'$, $\Tfr \RewContextRT \Tfr'$ imply $\Tfr' = \Tfr$.
In other words, a normal form for $\RewContext$ is a tree which is not
rewritable by $\RewContext$. The set of all the normal forms of
$\RewContext$ is denoted by $\NormalForms_{\RewContext}$.
\medbreak

\begin{Lemma} \label{lem:normal_forms_avoiding}
    Let $\Rew$ be a set of rewrite rules on $\Mag$ and $\RewContext$ be
    the rewrite relation induced by $\Rew$. Then, if $\RewContext$ is
    terminating, $\NormalForms_{\RewContext}$ is the set of all the
    binary trees that avoid all the trees appearing as left members
    of $\Rew$.
\end{Lemma}
\begin{proof}
    \Todo{à faire}
\end{proof}
\medbreak

When for any binary trees $\Tfr$, $\Sfr_1$, and $\Sfr_2$ such that
$\Tfr \RewContextRT \Sfr_1$ and $\Tfr \RewContextRT \Sfr_2$,
there exists a binary tree $\Tfr'$ such that
$\Sfr_1 \RewContextRT \Tfr'$ and $\Sfr_2 \RewContextRT \Tfr'$,
we say that $\RewContext$ is \Def{convergent}.

We use here the standard terminology
(\Def{terminating}, \Def{confluent}, \Def{convergent}, \Def{branching
pair}, \Def{joinable}, \Def{normal form}, {\em etc.}) about rewrite
relations and rewrite systems~\cite{BN98}.
\medbreak

Moreover, we shall use the following result appearing in~\cite{Gir16}
specialized on rewrite relation on $\Mag$ to prove that a terminating
rewrite relation is convergent.
\medbreak

\begin{Lemma} \label{lem:degree_confluence}
    Let $\Rew$ be a rewrite rule on $\Mag$ wherein all trees $\Tfr$ and
    $\Tfr'$ such that $\Tfr \Rew \Tfr'$ have degrees at most $\ell$.
    Then, if the rewrite relation $\RewContext$ induced by $\Rew$ is
    terminating and all its branching pairs of degrees at most
    $2\ell - 1$ are joinable, $\RewContext$ is convergent.
\end{Lemma}
\medbreak

Given an operad $\Oca \simeq \Mag/_{\Congr}$ where $\Congr$ is an operad
congruence of $\Mag$, we denote by $[\Tfr]_{\Congr}$ the
$\Congr$-equivalence class of any $\Tfr \in \Mag$. Besides, we say that
$\Rew$ is an \Def{orientation} of $\Congr$ if the reflexive, transitive,
and symmetric closure of $\RewContext$ is $\Congr$. We say that $\Rew$
is a \Def{convergent orientation} if $\RewContext$ is convergent. When
$\Rew$ is a convergent orientation of $\Congr$, the set of all normal
forms of $\RewContext$ is a \Def{Poincaré-Birkhoff-Witt basis} of the
operad $\Oca$ and its elements are exactly the binary trees avoiding,
as subtrees, the trees appearing as left members in~$\Rew$.
\medbreak