### 2018-08-28 CAs_operads.tex
Lemme 3.2.6

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Generalizations of the associative operad}
\label{sec:CAs_d}

In this section, we define comb associative operads and we show that the
set of such operads admits a lattice structure, isomorphic to the lattice
division for nonnegative integers. We relate this lattice to the
lattice of linear magmatic quotients. We also provide a finite convergent
orientation for the comb associative operad corresponding to $3$.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comb associative operads}
Recall first that the \Def{associative operad} $\As$ is the quotient
of $\Mag$ by the smallest operad congruence $\Congr$ satisfying
\begin{equation}
    \begin{tikzpicture}[xscale=.24,yscale=.24,Centering]
        \node(0)at(0.00,-3.33){};
        \node(2)at(2.00,-3.33){};
        \node(4)at(4.00,-1.67){};
        \node[NodeST](1)at(1.00,-1.67)
            {\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00)
            {\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(3);
        \node(r)at(3.00,1.5){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}
    \Congr
    \begin{tikzpicture}[xscale=.24,yscale=.24,Centering]
        \node(0)at(0.00,-1.67){};
        \node(2)at(2.00,-3.33){};
        \node(4)at(4.00,-3.33){};
        \node[NodeST](1)at(1.00,0.00)
                {\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.67)
                {\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \node(r)at(1.00,1.5){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}\,.
\end{equation}
We propose here a generalization of $\Congr$ in order to define
generalizations of $\As$.
\medbreak

Let, for any integer $\gamma \geq 1$, the binary trees $\LComb{\gamma}$
and $\RComb{\gamma}$ be respectively the left and the right comb of
degree $\gamma$. These trees are depicted as
\begin{equation}
    \LComb{\gamma} = \enspace
    \begin{tikzpicture}[xscale=.26,yscale=.3,Centering]
        \node(0)at(0.00,-5.25){};
        \node(2)at(2.00,-5.25){};
        \node(4)at(4.00,-3.50){};
        \node(6)at(6.00,-1.75){};
        \node[NodeST](1)at(1.00,-3.50){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge,dotted](1)edge[]node[font=\tiny]{
            \begin{math}\gamma\! -\! 1\end{math}\hspace*{.6cm}}(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(5.00,1.5){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \qquad \mbox{and} \qquad
    \RComb{\gamma} =
    \begin{tikzpicture}[xscale=.26,yscale=.3,Centering]
        \node(0)at(0.00,-1.75){};
        \node(2)at(2.00,-3.50){};
        \node(4)at(4.00,-5.25){};
        \node(6)at(6.00,-5.25){};
        \node[NodeST](1)at(1.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-3.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge,dotted](5)edge[]node[font=\tiny]{
            \hspace*{.6cm}\begin{math}\gamma\! -\! 1\end{math}}(3);
        \draw[Edge](6)--(5);
        \node(r)at(1.00,1.5){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}\,,
\end{equation}
where the values on the dotted edges denote the number of internal
nodes they contain. We shall employ this drawing convention in the
sequel, combined with the convention stipulating that dotted edges
with no value have any number of internal nodes. Let us now define for
any $\gamma \geq 1$ the \Def{$\gamma$-comb associative operad}
$\CAs{\gamma}$ as the quotient operad $\Mag/_{\CongrCAs{\gamma}}$ where
$\CongrCAs{\gamma}$ is the smallest operad congruence of $\Mag$
satisfying
\begin{equation} \label{equ:congruence_CAs_gamma}
    \LComb{\gamma} \enspace \CongrCAs{\gamma} \enspace \RComb{\gamma}.
\end{equation}
Notice that $\CongrCAs{1}$ is trivial so that
$\CAs{1} = \Mag$ and that $\CongrCAs{2}$ is the operad congruence
defining $\As$ so that $\CAs{2} = \As$. Let also
\begin{equation}
    \CAsAll := \left\{\CAs{\gamma} : \gamma\geq 1\right\}
\end{equation}
be the set of all the $\gamma$-comb associative operads.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Lattice of comb associative operads}

In order to introduce the lattice structure of $\CAsAll$, we study operad
morphisms betweeen its elements by mean of intermediate lemmas.
\medbreak

\begin{Lemma} \label{lem:first_dimensions_CAs}
    For all positive integers $\gamma$ and $n$ such that $\gamma \geq 2$
    and $n \leq \gamma + 1$,
    \begin{equation}
        \# \CAs{\gamma}(n) =
        \begin{cases}
            \Catalan(n)
                & \mbox{ if } n \leq \gamma, \\
            \Catalan(\gamma + 1) - 1
                & \mbox{ otherwise } (n = \gamma + 1).
        \end{cases}
    \end{equation}
\end{Lemma}

\begin{proof}
    Since the equivalence relation $\CongrCAs{\gamma}$ is trivial on the
    binary trees of degrees $d < \gamma$, and since a binary tree of
    degree $d$ has arity $n := d + 1$, one has
    $\# \CAs{\gamma}(n) = \# \Mag(n) = \Catalan(n)$ with $n \leq \gamma$.
    Besides, by definition of $\CongrCAs{\gamma}$, all the
    $\CongrCAs{\gamma}$-equivalence classes of binary trees of degree
    $\gamma$ are trivial, except one consisting in the pair
    $\left\{\LComb{\gamma}, \RComb{\gamma}\right\}$. Therefore, since a
    binary tree of degree $d$ has arity $n := \gamma + 1$,
    \begin{math}
        \# \CAs{\gamma}(n)
        = \# \Mag(\gamma + 1) - 1
        = \Catalan(\gamma + 1) - 1.
    \end{math}
\end{proof}
\medbreak

\begin{Lemma} \label{lem:surjective_morphisms_CAs}
    Let $\gamma$ and $\gamma'$ be two positive integers. If
    $\varphi:\CAs{\gamma'} \to \CAs{\gamma}$ is an operad morphism, then
    it is surjective and is unique.
\end{Lemma}

\begin{proof}
    The operad $\CAs{\gamma'}$ is generated by one binary generator
    $[\Product]_{\CongrCAs{\gamma'}}$, which is the image of $\Product$
    in $\CAs{\gamma'}$. Hence, $\varphi$ is entirely determined by the
    image $\varphi\left([\Product]_{\CongrCAs{\gamma'}}\right)$.
    Moreover, $\varphi([\Product]_{\CongrCAs{\gamma'}})$ has to be of
    arity $2$ in $\CAs{\gamma}$, so that we necessarily have
    \begin{math}
        \varphi\left([\Product]_{\CongrCAs{\gamma'}}\right)
        =
        [\Product]_{\CongrCAs{\gamma}}.
    \end{math}
    Hence, if $\varphi$ exists, it is the unique operad morphism from
    $\CAs{\gamma'}$ to $\CAs{\gamma}$. In this case,
    $[\Product]_{\CongrCAs{\gamma'}}$ being in the image of $\varphi$,
    the latter is surjective.
\end{proof}
\medbreak

\begin{Lemma} \label{lem:injective_morphisms_CAs}
    Let $\gamma$ and $\gamma'$ be two positive integers and
    $\varphi:\CAs{\gamma'} \to \CAs{\gamma}$ be an operad morphism.
    Then, $\varphi$ is injective if and only if $\gamma = \gamma'$.
\end{Lemma}

\begin{proof}
    Assume that $\varphi$ is injective. By
    Lemma~\ref{lem:surjective_morphisms_CAs}, $\varphi$ is also
    surjective, so that $\varphi$ is an isomorphism. If
    $\gamma \ne \gamma'$, by Lemma~\ref{lem:first_dimensions_CAs}, there
    is a positive integer $n$ such that
    $\# \CAs{\gamma}(n) \ne \# \CAs{\gamma'}(n)$. This is contradictory
    with the fact that $\CAs{\gamma}$ and $\CAs{\gamma'}$ are
    isomorphic. Hence, $\gamma = \gamma'$.
    \smallbreak

    Conversely, if $\gamma = \gamma'$, the only operad morphism from
    $\CAs{\gamma}$ to itself sends the generator
    $[\Product]_{\CongrCAs{\gamma}}$ to itself. This maps extends as
    an operad morphism into the identity morphism which is of course
    injective.
\end{proof}
\medbreak

\begin{Lemma} \label{lem:morphism_CAs}
    Let $\gamma$ and $\gamma'$ be two positive integers. There exists an
    operad morphism $\varphi:\CAs{\gamma'} \to \CAs{\gamma}$ if and only
    if
    \begin{math}
      \LComb{\gamma'} \CongrCAs{\gamma} \RComb{\gamma'}.
    \end{math}
\end{Lemma}

\begin{proof}
    Assume that $\varphi:\CAs{\gamma'} \to \CAs{\gamma}$ is an operad
    morphism. By Lemma~\ref{lem:surjective_morphisms_CAs}, $\varphi$
    satisfies
    \begin{math}
        \varphi\left(\left[\Tfr\right]_{\CongrCAs{\gamma'}}\right)
        = \left[\Tfr\right]_{\CongrCAs{\gamma}}
    \end{math}
    for any binary tree $\Tfr$.
    Since
    \begin{math}
        \varphi\left(
        \left[\LComb{\gamma'}\right]_{\CongrCAs{\gamma'}}
        \right)
        =
        \varphi\left(
        \left[\RComb{\gamma'}\right]_{\CongrCAs{\gamma'}}
        \right),
    \end{math}
    we have
    \begin{math}
        \left[\LComb{\gamma'}\right]_{\CongrCAs{\gamma}}
        =
        \left[\RComb{\gamma'}\right]_{\CongrCAs{\gamma}},
    \end{math}
    which is equivalent to the fact that
    \begin{math}
        \LComb{\gamma'} \CongrCAs{\gamma} \RComb{\gamma'}.
    \end{math}
    \smallbreak

    Conversely, when
    \begin{math}
        \LComb{\gamma'}\CongrCAs{\gamma}\RComb{\gamma'},
    \end{math}
    let $\varphi:\CAs{\gamma'}(2) \to \CAs{\gamma}(2)$ be the map
    defined by
    \begin{math}
        \varphi\left(
        \left[\Product\right]_{\CongrCAs{\gamma'}}\right)
        :=
        \left[\Product\right]_{\CongrCAs{\gamma}}.
    \end{math}
    Now, since $\CongrCAs{\gamma}$ is coarser than $\CongrCAs{\gamma'}$,
    $\varphi$ extends (in a unique way) into an operad morphism, whence
    the statement of the lemma.
\end{proof}
\medbreak

We define the binary relation $\OrdCAs$ on $\CAsAll$ as follows: we have
$\CAs{\gamma} \OrdCAs \CAs{\gamma'}$ if and only if there exists a
morphism $\varphi:\CAs{\gamma'} \to \CAs{\gamma}$.
\medbreak

\begin{Proposition}\label{prop:poset_CAs}
    The binary relation $\OrdCAs$ is a partial order relation on
    $\CAsAll$.
\end{Proposition}

\begin{proof}
    The binary relation $\OrdCAs$ is reflexive since there exists the
    identity morphism of $\CAs{\gamma}$ for every positive integer
    $\gamma$. Let us now assume that
    $\varphi:\CAs{\gamma''} \to \CAs{\gamma'}$ and
    $\psi:\CAs{\gamma'} \to \CAs{\gamma}$ are two operad morphisms. By
    Lemma~\ref{lem:morphism_CAs}, one has
    $\LComb{\gamma''} \CongrCAs{\gamma'} \RComb{\gamma''}$ and
    $\LComb{\gamma'} \CongrCAs{\gamma} \RComb{\gamma'}$. This implies
    that $\CongrCAs{\gamma}$ is coarser than $\CongrCAs{\gamma'}$ and
    that $\CongrCAs{\gamma'}$ is coarser that $\CongrCAs{\gamma''}$. Now,
    again by Lemma~\ref{lem:morphism_CAs}, $\psi \circ \varphi$ is an
    operad morphism. Hence, $\OrdCAs$ is transitive. Finally, let us
    assume that there exist morphisms
    $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$ and
    $\psi:\CAs{\gamma}\to\CAs{\gamma'}$. In particular,
    $\psi \circ \varphi$ and $\varphi \circ \psi$ are endomorphisms of
    $\CAs{\gamma'}$ and $\CAs{\gamma}$, respectively. From
    Lemma~\ref{lem:surjective_morphisms_CAs}, these two morphisms are
    identity morphisms, so that $\varphi$ and $\psi$ are injective. From
    Lemma~\ref{lem:injective_morphisms_CAs}, $\gamma$ and $\gamma'$ are
    equal, which proves that $\OrdCAs$ is anti-symmetric. Hence,
    $\OrdCAs$ is a partial order.
\end{proof}
\medbreak

In order to show that $\left(\CAsAll, \OrdCAs\right)$ extends into a
lattice, we relate $\left(\CAsAll, \OrdCAs\right)$ with the lattice of
integers $\left(\N,\mid, \gcd, \lcm\right)$, where $\mid$ denotes the
division relation, $\gcd$ denotes the greatest common divisor, and
$\lcm$ the least common multiple operators, respectively.

Recall that
$\LRank(\Tfr)$ denotes the left rank of a binary tree $\Tfr$, defined in
Section~\ref{sec:operad_Mag}. Besides, to simplify the notation,
we shall write $\bar{a}$ instead of $a - 1$ for any integer $a$.
\medbreak

\begin{Lemma}\label{lem:left_rank_and_CongrCAs}
    Let $\gamma$ be a positive integer and let $\Tfr$ and
    $\Tfr'$ be two binary trees. If $\Tfr \CongrCAs{\gamma} \Tfr'$, then
    \begin{equation}
        \LRank(\Tfr) \pmod{\bar{\gamma}}
        \enspace = \enspace
        \LRank\left(\Tfr'\right) \pmod{\bar{\gamma}}.
    \end{equation}
\end{Lemma}

\Todo{Christophe : je vous propose cette démo}
\begin{proof}
    We consider the rewrite rule $\LComb{\gamma} \Rew \RComb{\gamma}$ 
    on $\Mag$. First, we show that 
    \begin{equation} \label{eq:rewriting_invariant}
        \Tfr \RewContext \Tfr' 
        \mbox{ implies that }
        \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right) 
        \, \pmod{\bar{\gamma}}.
    \end{equation}
    Indeed for all $\Tfr \in \Mag$, there exist binary trees 
    $\Tfr_1, \dots, \Tfr_{\LRank\left(\Tfr\right)}$ such that
    \begin{math}
      \Tfr = 
      \left[\, |,\Tfr_1, \dots, \Tfr_{\LRank\left(\Tfr\right)} \right].
    \end{math}
    Thus if $\Tfr \RewContext \Tfr'$, then one of the 
    two following cases occurs :
    \begin{enumerate}[label={(\it\roman*)}]
        \item there exist $\Tfr_i'$ such that
            \begin{math}
              \Tfr' = \LComb{\LRank\left(\Tfr\right)} \circ 
              \left[\, | ,\Tfr_1, \dots, \Tfr_i',\dots, 
                    \Tfr_{\LRank\left(\Tfr\right)}
              \right] \mbox{ so } 
              \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right);
            \end{math}
        \item there exist $i$ such that
            \begin{math} 
              \Tfr' = \LComb{\LRank\left(\Tfr\right)-\bar{\gamma}} \circ
              \left[ | ,\Tfr_1, \dots, 
                 \RComb{\bar{\gamma}} \circ 
                 [\Tfr_{i},\dots, \Tfr_{i+\bar{\gamma}}]
                 ,\dots, \Tfr_{\LRank\left(\Tfr\right)}
              \right]\\
              \mbox{ so }
              \LRank\left(\Tfr'\right) = 
              \LRank\left(\Tfr\right) - \bar{\gamma} = 
              \LRank\left(\Tfr\right) \pmod{\bar{\gamma}}
            \end{math}.
    \end{enumerate} 
    This concludes the proof of~\ref{eq:rewriting_invariant}.  
    \smallbreak

    We deduce from~\ref{eq:rewriting_invariant} that the equivalence 
    relation $r$ defined by 
    \begin{equation}
        r(\Tfr, \Tfr') \mbox{ if and only if } 
        \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right) 
        \, \pmod{\bar{\gamma}}
    \end{equation}
    contains the relation $\RewContext$ and thus its transitive 
    and symmetric closure, which is $\CongrCAs{\gamma}$.
\end{proof}

\begin{proof}
  We consider the rewrite rule $\LComb{\gamma} \Rew \RComb{\gamma}$ on
  $\Mag$.
    \smallbreak

    First, we show that $\Tfr \RewContext \Tfr'$ implies that
    $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by $\bar{\gamma}$. \Todo{à faire.}
    \smallbreak

    Assume that $\Tfr \RewContextRT \Tfr'$, so that there exist binary trees
    $\Tfr_1$, \dots, $\Tfr_k$ such that $\Tfr_1 = \Tfr$,
    $\Tfr_k = \Tfr'$ and for every $i \in [k - 1]$, we have
    $\Tfr_i\RewContext \Tfr_{i + 1}$. The integer
    \begin{math}
        \LRank\left(\Tfr_{i + 1}\right)
        -
        \LRank\left(\Tfr_i\right)
    \end{math}
    is divisible by $\bar{\gamma}$ from the first part of the proof, so
    that $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by~$\bar{\gamma}$.
    \smallbreak

    Assume that $\Tfr \RewContextRST \Tfr'$, so that there exist binary trees
    $\Tfr_1$, \dots, $\Tfr_k$ such that $\Tfr_1 = \Tfr$,
    $\Tfr_k = \Tfr'$ and for every $i \in [k - 1]$, we have
    $\Tfr_i \RewContextRT \Tfr_{i + 1}$ or $\Tfr_{i + 1} \RewContextRT
    \Tfr_i$. From the second part of the proof, the integer
    \begin{math}
        \LRank\left(\Tfr_{i + 1}\right)
        -
        \LRank\left(\Tfr_i\right)
    \end{math}
    is divisible by $\bar{\gamma}$, so that
    $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by~$\bar{\gamma}$.
    \smallbreak

    The congruence $\CongrCAs{\gamma}$ is the equivalence relation
    induced by the rewrite relation $\RewContext$ induced by the rewrite
    rule $\LComb{\gamma} \Rew \RComb{\gamma}$, that is we have $\Tfr
    \CongrCAs{\gamma} \Tfr'$ if and only if $\Tfr \RewContextRST \Tfr'$.
    Hence, from the third part of the proof, $\Tfr \CongrCAs{\gamma}
    \Tfr'$ implies that
    \begin{math}
        \LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)
    \end{math}
    is divisible by~$\bar{\gamma}$.
\end{proof}
\medbreak

\begin{Proposition} \label{prop:division_CAs}
    Let $\gamma$ and $\gamma'$ be two positive integers. Then, there
    exists a morphism $\varphi:\CAs{\gamma'} \to \CAs{\gamma}$ if and
    only if $\bar{\gamma} \mid \bar{\gamma'}$.
\end{Proposition}

\begin{proof}
    From Lemma~\ref{lem:morphism_CAs}, it is enough to show that
    $\LComb{\gamma'} \CongrCAs{\gamma} \RComb{\gamma'}$ if and only if
    $\bar{\gamma} \mid \bar{\gamma'}$. If
    $\LComb{\gamma'}\CongrCAs{\gamma}\RComb{\gamma'}$, from
    Lemma~\ref{lem:left_rank_and_CongrCAs},
    \begin{math}
        \LRank\left(\LComb{\gamma'}\right)
        - \LRank\left(\RComb{\gamma'}\right)
        = \bar{\gamma'}
    \end{math}
    is divisible by $\bar{\gamma}$, which shows the direct implication.
    \smallbreak

    Conversely, if
    $\bar{\gamma} \mid \bar{\gamma'}$, the rewrite
    rule $\LComb{\gamma} \Rew \RComb{\gamma}$ induces the
    sequence
    \begin{multline}
        \LComb{\gamma'} \enspace = \enspace
        \begin{tikzpicture}[xscale=.19,yscale=.25,Centering]
            \node(0)at(0.00,-13.12){};
            \node(10)at(10.00,-5.62){};
            \node(12)at(12.00,-3.75){};
            \node(14)at(14.00,-1.88){};
            \node(2)at(2.00,-13.12){};
            \node(4)at(4.00,-11.25){};
            \node(6)at(6.00,-9.38){};
            \node(8)at(8.00,-7.50){};
            \node[NodeST](1)at(1.00,-11.25)
                {\begin{math}\Product\end{math}};
            \node[NodeST](11)at(11.00,-1.88)
                {\begin{math}\Product\end{math}};
            \node[NodeST](13)at(13.00,0.00)
                {\begin{math}\Product\end{math}};
            \node[NodeST](3)at(3.00,-9.38)
                {\begin{math}\Product\end{math}};
            \node[NodeST](5)at(5.00,-7.50)
                {\begin{math}\Product\end{math}};
            \node[NodeST](7)at(7.00,-5.62)
                {\begin{math}\Product\end{math}};
            \node[NodeST](9)at(9.00,-3.75)
                {\begin{math}\Product\end{math}};
            \draw[Edge](0)--(1);
            \draw[Edge,dotted](1)edge[]node[font=\footnotesize]{
                \begin{math}\bar{\gamma}\end{math}\hspace*{.5cm}}(3);
            \draw[Edge](10)--(9);
            \draw[Edge](11)--(13);
            \draw[Edge](12)--(11);
            \draw[Edge](14)--(13);
            \draw[Edge](2)--(1);
            \draw[Edge,dotted](3)--(5);
            \draw[Edge](4)--(3);
            \draw[Edge,dotted](5)edge[]node[font=\footnotesize]{
                \begin{math}\bar{\gamma}\end{math}\hspace*{.5cm}}(7);
            \draw[Edge](6)--(5);
            \draw[Edge](7)--(9);
            \draw[Edge](8)--(7);
            \draw[Edge,dotted](9)edge[]node[font=\footnotesize]{
                \begin{math}\bar{\gamma}\end{math}\hspace*{.5cm}}(11);
            \node(r)at(13.00,1.41){};
            \draw[Edge](r)--(13);
            %
            %\node[fit=(1)(3),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
            %\node[fit=(5)(7),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
            %\node[fit=(9)(11),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
        \end{tikzpicture}
        \enspace \RewContext \enspace
        \begin{tikzpicture}[xscale=.19,yscale=.2,Centering]
            \node(0)at(0.00,-12.50){};
            \node(10)at(10.00,-5.00){};
            \node(12)at(12.00,-7.50){};
            \node(14)at(14.00,-7.50){};
            \node(2)at(2.00,-12.50){};
            \node(4)at(4.00,-10.00){};
            \node(6)at(6.00,-7.50){};
            \node(8)at(8.00,-5.00){};
            \node[NodeST](1)at(1.00,-10.00)
                {\begin{math}\Product\end{math}};
            \node[NodeST](11)at(11.00,-2.50)
                {\begin{math}\Product\end{math}};
            \node[NodeST](13)at(13.00,-5.00)
                {\begin{math}\Product\end{math}};
            \node[NodeST](3)at(3.00,-7.50)
                {\begin{math}\Product\end{math}};
            \node[NodeST](5)at(5.00,-5.00)
                {\begin{math}\Product\end{math}};
            \node[NodeST](7)at(7.00,-2.50)
                {\begin{math}\Product\end{math}};
            \node[NodeST](9)at(9.00,0.00)
                {\begin{math}\Product\end{math}};
            \draw[Edge](0)--(1);
            \draw[Edge,dotted](1)edge[]node[font=\footnotesize]{
                \begin{math}\bar{\gamma}\end{math}\hspace*{.5cm}}(3);
            \draw[Edge](10)--(11);
            \draw[Edge](11)--(9);
            \draw[Edge](12)--(13);
            \draw[Edge,dotted](13)edge[]node[font=\footnotesize]{
                \hspace*{.5cm}\begin{math}\bar{\gamma}\end{math}}(11);
            \draw[Edge](14)--(13);
            \draw[Edge](2)--(1);
            \draw[Edge,dotted](3)--(5);
            \draw[Edge](4)--(3);
            \draw[Edge,dotted](5)edge[]node[font=\footnotesize]{
                \begin{math}\bar{\gamma}\end{math}\hspace*{.5cm}}(7);
            \draw[Edge](6)--(5);
            \draw[Edge](7)--(9);
            \draw[Edge](8)--(7);
            \node(r)at(9.00,1.88){};
            \draw[Edge](r)--(9);
            %
            %\node[fit=(1)(3),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
            %\node[fit=(5)(7),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
            %\node[fit=(11)(13),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
        \end{tikzpicture} \\
        \enspace \RewContext \enspace
        \begin{tikzpicture}[xscale=.19,yscale=.2,Centering]
            \node(0)at(0.00,-7.50){};
            \node(10)at(10.00,-10.00){};
            \node(12)at(12.00,-12.50){};
            \node(14)at(14.00,-12.50){};
            \node(2)at(2.00,-7.50){};
            \node(4)at(4.00,-5.00){};
            \node(6)at(6.00,-5.00){};
            \node(8)at(8.00,-7.50){};
            \node[NodeST](1)at(1.00,-5.00)
                {\begin{math}\Product\end{math}};
            \node[NodeST](11)at(11.00,-7.50)
                {\begin{math}\Product\end{math}};
            \node[NodeST](13)at(13.00,-10.00)
                {\begin{math}\Product\end{math}};
            \node[NodeST](3)at(3.00,-2.50)
                {\begin{math}\Product\end{math}};
            \node[NodeST](5)at(5.00,0.00)
                {\begin{math}\Product\end{math}};
            \node[NodeST](7)at(7.00,-2.50)
                {\begin{math}\Product\end{math}};
            \node[NodeST](9)at(9.00,-5.00)
                {\begin{math}\Product\end{math}};
            \draw[Edge](0)--(1);
            \draw[Edge,dotted](1)edge[]node[font=\footnotesize]{
                \begin{math}\bar{\gamma}\end{math}\hspace*{.5cm}}(3);
            \draw[Edge](10)--(11);
            \draw[Edge](11)--(9);
            \draw[Edge](12)--(13);
            \draw[Edge,dotted](13)edge[]node[font=\footnotesize]{
                \hspace*{.5cm}\begin{math}\bar{\gamma}\end{math}}(11);
            \draw[Edge](14)--(13);
            \draw[Edge](2)--(1);
            \draw[Edge,dotted](3)--(5);
            \draw[Edge](4)--(3);
            \draw[Edge](6)--(7);
            \draw[Edge](7)--(5);
            \draw[Edge](8)--(9);
            \draw[Edge,dotted](9)edge[]node[font=\footnotesize]{
                \hspace*{.5cm}\begin{math}\bar{\gamma}\end{math}}(7);
            \node(r)at(5.00,1.88){};
            \draw[Edge](r)--(5);
            %
            %\node[fit=(1)(3),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
            %\node[fit=(7)(9),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
            %\node[fit=(11)(13),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
        \end{tikzpicture}
        \enspace \RewContextRT \enspace
        \begin{tikzpicture}[xscale=.19,yscale=.25,Centering]
            \node(0)at(0.00,-1.88){};
            \node(10)at(10.00,-11.25){};
            \node(12)at(12.00,-13.12){};
            \node(14)at(14.00,-13.12){};
            \node(2)at(2.00,-3.75){};
            \node(4)at(4.00,-5.62){};
            \node(6)at(6.00,-7.50){};
            \node(8)at(8.00,-9.38){};
            \node[NodeST](1)at(1.00,0.00)
                {\begin{math}\Product\end{math}};
            \node[NodeST](11)at(11.00,-9.38)
                {\begin{math}\Product\end{math}};
            \node[NodeST](13)at(13.00,-11.25)
                {\begin{math}\Product\end{math}};
            \node[NodeST](3)at(3.00,-1.88)
                {\begin{math}\Product\end{math}};
            \node[NodeST](5)at(5.00,-3.75)
                {\begin{math}\Product\end{math}};
            \node[NodeST](7)at(7.00,-5.62)
                {\begin{math}\Product\end{math}};
            \node[NodeST](9)at(9.00,-7.50)
                {\begin{math}\Product\end{math}};
            \draw[Edge](0)--(1);
            \draw[Edge](10)--(11);
            \draw[Edge,dotted](11)--(9);
            \draw[Edge](12)--(13);
            \draw[Edge,dotted](13)edge[]node[font=\footnotesize]{
                \hspace*{.5cm}\begin{math}\bar{\gamma}\end{math}}(11);
            \draw[Edge](14)--(13);
            \draw[Edge](2)--(3);
            \draw[Edge](3)--(1);
            \draw[Edge](4)--(5);
            \draw[Edge,dotted](5)edge[]node[font=\footnotesize]{
                \hspace*{.5cm}\begin{math}\bar{\gamma}\end{math}}(3);
            \draw[Edge](6)--(7);
            \draw[Edge](7)--(5);
            \draw[Edge](8)--(9);
            \draw[Edge,dotted](9)edge[]node[font=\footnotesize]{
                \hspace*{.5cm}\begin{math}\bar{\gamma}\end{math}}(7);
            \node(r)at(1.00,1.41){};
            \draw[Edge](r)--(1);
            %
            %\node[fit=(3)(5),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
            %\node[fit=(7)(9),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
            %\node[fit=(11)(13),rotate fit=0,inner sep=0pt,
                %rounded corners,draw=Col5!80]{};
        \end{tikzpicture}
        \enspace = \enspace
        \RComb{\gamma'}
    \end{multline}
    of rewriting steps, where squared regions denotes left or right
    comb trees of degree $\gamma - 1$. Hence, we have
    $\LComb{\gamma'}\CongrCAs{\gamma}\RComb{\gamma'}$.
\end{proof}
\medbreak

Proposition~\ref{prop:division_CAs} implies the following result.
\medbreak

\begin{Theorem}\label{thm:lattice_CAs}
  The tuple $\LatCAs$ is a lattice, where $\InfCAs$ and $\SupCAs$ are
  defined, for all positive integers $\gamma$ and $\gamma'$, by
    \begin{equation}
      \CAs{\gamma}\InfCAs\CAs{\gamma'}
      =\CAs{\gcd \left(\bar{\gamma}, \bar{\gamma'}\right)}
    \end{equation}
    and
    \begin{equation}
      \CAs{\gamma}\SupCAs\CAs{\gamma'}
      =\CAs{\lcm \left(\bar{\gamma}, \bar{\gamma'}\right)}.
    \end{equation}
\end{Theorem}
\medbreak

The lattice $\left(\N,\mid, \gcd, \lcm\right)$ admits $1$ as minimum
element and $0$ as maximum element: any nonnegative integer is
divisible by $1$ and divides $0$. The corresponding property for
$\LatCAs$ is that it admits $\As=\CAs{2}$ as minimum element and $\Mag=
\CAs{1}$ as maximum element: any comb associative operad projects onto
$\As$ and is a quotient of $\Mag$.

We terminate this section by relating the lattice
$\LatQMag$ introduced in Section~\ref{sec:Magmatic_operads} to the
lattice $\LatCAs$.

As explained in Section~\ref{sec:operad_Mag}, a set-theoreic operad can
be embedded into a linear operad, so that the operads $\CAs{\gamma}$ can
be embedded into quotient operads $\KCAs{\gamma}$ of $\KMag$. Formally,
the operad $\KCAs{\gamma}$ is equal to $\KMag/_{I_{\gamma}}$, where
$I_{\gamma}$ is the operadic ideal of $\KMag$ generated by
$\LComb{\gamma}-\RComb{\gamma}$.

We obtain a new lattice $\LatKCAs$, where $\KCAsAll$ is the set of all
operads $\KCAs{\gamma}$. In this linear framework, the condition
$\KCAs{\gamma}\OrdCAs\KCAs{\gamma'}$ means that the dimension of the
space $\text{hom}\left(\KCAs{\gamma'},\KCAs{\gamma}\right)$ is equal to
$1$. Hence, $\LatKCAs$ is related to $\LatQMag$ by the following
theorem.

\begin{Theorem} \label{thm:inclusion_lattice_CAs}
    The inclusion
    \begin{math}
        \iota:\left(\KCAsAll,\OrdCAs\right)
        \to
        \left(\QMag,\OrdQMag\right)
    \end{math}
    is nondecreasing. In particular, for all positive integers $\gamma$
    and $\gamma'$, we have
    \begin{equation} \label{equ:comparison_of_lattice_operations}
        \KCAs{\gcd \left(\bar{\gamma}, \bar{\gamma'}\right)}
        \OrdQMag
        \KCAs{\gamma}\InfQMag\KCAs{\gamma'}
    \end{equation}
    and
    \begin{equation}
        \KCAs{\gamma}\SupQMag\KCAs{\gamma'}
        \OrdQMag
        \KCAs{\lcm \left(\bar{\gamma}, \bar{\gamma'}\right)}.
    \end{equation}
\end{Theorem}
\medbreak

Note that $\LatKCAs$ does not embed as a sub-lattice of $\LatQMag$, that
is $\iota$ is not a lattice morphism. Consider for instance $\gamma = 3$
and $\gamma' = 4$, then $\KCAs{\gamma} \InfCAs \KCAs{\gamma'}$ is equal
to $\KCAs{2} = \K\As$, whereas $\KCAs{\gamma} \InfQMag \KCAs{\gamma'}$
is a quotient of $\KMag$ be the two relations $\LComb{3} - \RComb{3}$
and $\LComb{4} - \RComb{4}$.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The \texorpdfstring{$3$}{3}-comb associative operad}
\label{subsec:CAs_3}
We now focus on the study of the operad $\CAs{3}$. By definition, this
operad is the quotient of $\Mag$ by the operad congruence spanned by the
relation
\begin{equation} \label{equ:rew_1}
    \begin{tikzpicture}[xscale=.22,yscale=.23,Centering]
        \node(0)at(0.00,-5.25){};
        \node(2)at(2.00,-5.25){};
        \node(4)at(4.00,-3.50){};
        \node(6)at(6.00,-1.75){};
        \node[NodeST](1)at(1.00,-3.50){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(5.00,1.5){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \enspace \Rew \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.23,Centering]
        \node(0)at(0.00,-1.75){};
        \node(2)at(2.00,-3.50){};
        \node(4)at(4.00,-5.25){};
        \node(6)at(6.00,-5.25){};
        \node[NodeST](1)at(1.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-3.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(1.00,1.5){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}\,.
\end{equation}
This rewrite rule is compatible with the lexicographic order on prefix
words presented at the beginning of Section~\ref{sec:operad_Mag} in the
sense that the prefix word of the left member of~\eqref{equ:rew_1} is
lexicographically greater than the prefix word of the right one.
\medbreak

However, the rewrite relation $\RewContext$ induced by $\Rew$ is not
confluent. Indeed, we have
\begin{equation} \label{equ:branching_pair_CAs_3}
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-7.20){};
        \node(2)at(2.00,-7.20){};
        \node(4)at(4.00,-5.40){};
        \node(6)at(6.00,-3.60){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](5)--(7);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.35){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \enspace \RewContext \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.20,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-4.50){};
        \node(4)at(4.00,-4.50){};
        \node(6)at(6.00,-6.75){};
        \node(8)at(8.00,-6.75){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-4.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(3.00,1.74){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}
    \qquad \mbox{and} \qquad
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-7.20){};
        \node(2)at(2.00,-7.20){};
        \node(4)at(4.00,-5.40){};
        \node(6)at(6.00,-3.60){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](5)--(7);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.35){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \enspace \RewContext \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.60){};
        \node(2)at(2.00,-5.40){};
        \node(4)at(4.00,-7.20){};
        \node(6)at(6.00,-7.20){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(7);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.5){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}\,,
\end{equation}
and the two right members of~\eqref{equ:branching_pair_CAs_3} form a
branching pair which is not joinable.
\medbreak

In order to transform the rewrite relation induced by~\eqref{equ:rew_1}
into a convergent one, we apply the Buchberger algorithm for
operads~\cite[Section 3.7]{DK10} with respect to the lexicographic order
on prefix words. Following this algorithm, we need to put the right
members of~\eqref{equ:branching_pair_CAs_3} in relation by $\Rew$. To
respect the lexicographic property of the prefix words, this leads to
the new relation
\begin{equation} \label{equ:rew_2}
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.60){};
        \node(2)at(2.00,-5.40){};
        \node(4)at(4.00,-7.20){};
        \node(6)at(6.00,-7.20){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(7);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.5){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \enspace \Rew \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.20,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-4.50){};
        \node(4)at(4.00,-4.50){};
        \node(6)at(6.00,-6.75){};
        \node(8)at(8.00,-6.75){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-4.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(3.00,1.74){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}\,.
\end{equation}
The Buchberger algorithm applied on binary trees of degrees $5$, $6$,
and $7$ provides the new relations \smallbreak
\input{Buchberger_relations_5-7}
\medbreak

\begin{Theorem} \label{thm:convergent_rewrite_rule_CAs_3}
    The set $\Rew$ of rewrite rules containing~\eqref{equ:rew_1},
    \eqref{equ:rew_2}, \eqref{equ:rew_3}---\eqref{equ:rew_11} is a
    convergent orientation of the congruence $\CongrCAs{3}$
    of~$\CAs{3}$.
\end{Theorem}
\begin{proof}
    Let us show that the rewrite relation $\RewContext$ induced by
    $\Rew$ is convergent. First, for every relation $\Tfr \Rew \Tfr'$,
    we have $\Tfr > \Tfr'$. Therefore, by
    Lemma~\ref{lem:prefix_word_termination}, $\RewContext$ is
    terminating. Moreover, the greatest degree of a tree appearing in
    $\Rew$ is~$7$ so that, from Lemma~\ref{lem:degree_confluence}, to
    show that $\RewContext$ is convergent, it is enough to prove that
    each tree of degree at most $13$ admits exactly one normal form.
    Equivalently, this amounts to show that the number of normal forms
    of trees of arity $n$ is equal to $\#\CAs{3}(n)$. By computer
    exploration, we get the same sequence
    \begin{equation} \label{equ:dimensions_CAs_3}
        1, 1, 2, 4, 8, 14, 20, 19, 16, 14, 14, 15, 16, 17
    \end{equation}
    for $\#\CAs{3}(n)$ and for the numbers of normal forms of arity $n$,
    when $ 1 \leq n \leq 14$. Hence, we get our following main result.
\end{proof}
\medbreak

The rewrite rule $\Rew$ has, arity by arity, the cardinalities
\begin{equation}
    0, 0, 0, 1, 1, 2, 3, 4, 0, 0, \dots~.
\end{equation}
We obtain from Theorem~\ref{thm:convergent_rewrite_rule_CAs_3} also
the following consequences.
\medbreak

\begin{Proposition} \label{prop:PBW_basis_CAs_3}
    The set of the trees avoiding as subtrees the ones appearing as
    left members of $\Rew$ is a Poincaré-Birkhoff-Witt basis
    of~$\CAs{3}$.
\end{Proposition}
\begin{proof}
    By definition of PBW bases and
    Theorem~\ref{thm:convergent_rewrite_rule_CAs_3}, the set
    $\NormalForms_{\RewContext}$ is a PBW basis of $\CAs{3}$ where
    $\RewContext$ is the rewrite relation induced by $\Rew$. Now, by
    Lemma~\ref{lem:normal_forms_avoiding}, $\NormalForms_{\RewContext}$
    can be described as the set of the trees avoiding the left members
    of~$\Rew$, whence the statement.
\end{proof}
\medbreak

\begin{Proposition} \label{prop:Hilbert_series_CAs_3}
    The Hilbert series of $\CAs{3}$ is
    \begin{equation} \label{equ:Hilbert_series_CAs_3}
        \HilbertSeries_{\CAs{3}}(t) = \frac{t}{(1 - t)^2}
        \left(1 - t + t^2 + t^3 + 2t^4 + 2t^5 - 7t^7 - 2t^8 + t^9 +
        2t^{10} + t^{11}\right).
    \end{equation}
\end{Proposition}
\begin{proof}
    From Proposition~\ref{prop:PBW_basis_CAs_3}, for any $n \geq 1$, the
    dimension of $\CAs{3}(n)$ is the number of trees that avoid as
    subtrees the left members of $\Rew$. Now, by using a result
    of~\cite{Gir18} (see also~\cite{Row10,KP15}) providing a system of
    equations for the generating series of the trees avoiding some sets
    of subtrees, we obtain Expression~\eqref{equ:Hilbert_series_CAs_3}
    for the considered family.
\end{proof}
\medbreak

For $n \leq 10$, the dimensions of $\CAs{3}(n)$ are provided by
Sequence~\eqref{equ:dimensions_CAs_3} and for all $n \geq 11$, the
Taylor expansion of~\eqref{equ:Hilbert_series_CAs_3} shows that
$\# \CAs{3}(n) = n + 3$.
\medbreak

\Todo{S: réalisation combinatoire de $\CAs{3}$ à donner ici.}

\Todo{S: résultats expérimentaux de quelques $\CAs{\gamma}$ à donner
dans cette partie (ou peut être dans une sous-partie précédente ; à
voir selon ce que ça donne).}

%By computer exploration, we have already some information in this
%direction: we have the sequence
%\begin{equation}
    %1, 1, 2, 5, 13, 35, 96, 264, 724, 1973, 5335, 14390, 38872, 105141,
    %284929, 774254
%\end{equation}
%for the first dimensions for $\CAs{4}$. By applying the Buchberger
%algorithm on trees of degrees until $10$, we obtain that a convergent
%orientation of $\CongrCAs{4}$ has, arity by arity, the sequence
%\begin{math}
    %0, 0, 0, 0, 1, 1, 0, 3, 4, 5, 18, 22
%\end{math}
%for its first cardinalities. Moreover, for $\CAs{5}$, we get the
%sequence
%\begin{equation}
%1, 1, 2, 5, 14, 41, 124, 384, 1210, 3861, 12440, 40392, 131997,
%433782, 1432696, 4752857
%\end{equation}
%of dimensions and the first cardinalities
%\begin{math}
    %0, 0, 0, 0, 0, 1, 1, 0, 0, 4, 5
%\end{math}
%for any convergent orientation of $\CongrCAs{5}$. Finally, for
%$\CAs{6}$, we get the sequence
%\begin{equation}
   %1, 1, 2, 5, 14, 42, 131, 420, 1375, 4576, 15431, 52598,
   %180895, 626862, 2186504, 7670138
%\end{equation}
%of dimensions and the first cardinalities
%\begin{math}
    %0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0
%\end{math}
%for any convergent orientation of $\CongrCAs{6}$. We can notice that
%only $\CAs{3}$ seems to have oscillating first dimensions.
%\medbreak