### 2018-09-10 Conclusion.tex
Relecture conclusion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusion and perspectives}
In this paper, we have considered some quotients of the magmatic operad
in both the linear and the set-theoretic frameworks. We focused mainly
our study on comb associative operads and collected properties by
using computer exploration and rewrite systems on trees. There are many
ways to extend this work. Here follow some few further research
directions.
\medbreak

A first research direction consists in finding convergent presentations
for (all or most of) the operads $\CAs{\gamma}$ in
order to describe algebraic and combinatorial properties of them (as
describing explicit bases, computing Hilbert series, and providing
combinatorial realizations). This has been done only in the case
$\gamma = 3$. For some other cases, we only have conjectural and
experimental data (see
Section~\ref{sec:higher_comb_associative_operads}).
\medbreak

Following ideas existing for word rewriting theory~\cite{GGM15}, a
second axis consists in allowing new generators for the operads
$\CAs{\gamma}$ in order to obtain finite convergent presentations
when $\gamma \geq 4$. Indeed, the Buchberger semi-algorithm works by
adding rewrite rules to a set of rewrite rules to obtain a convergent
rewrite system. An orthogonal procedure consists rather in adding new
generators (new labels for internal nodes in the trees) in order to
obtain convergent rewrite systems. More generally, we also would like to
use these ideas for other magmatic quotients, such as the operad
$\Mag^{\{3, 4\}}$ that we did not study entirely in
Section~\ref{sec:MAg_3}.
\medbreak

A third axis consists in studying if the completion of presentations of
quotients of the magmatic operad maintains links with
the lattice structure introduced in Section~\ref{sec:Magmatic_operads}.
More precisely, assuming that we have completed the presentations of
the quotients $\Oca_1$ and $\Oca_2$ of $\Mag$,
as well as the one of the lower-bound $\Oca_1 \InfQMag \Oca_2$, the
question consists in designing an algorithm for computing a completion
of a presentation of the upper-bound
$\Oca_1 \SupQMag \Oca_2$. Of course, the same question also makes sense
for the lattice of comb associative operads introduced in
Section~\ref{sec:CAs_d}.
\medbreak

Let us address now a perspective fitting more in a combinatorial
context. As mentioned in the introduction of this article, we suspect
that some combinatorial properties of quotients $\Mag/_{\Congr}$ of
$\Mag$ derive from properties of the equivalence relation generating the
operad congruence $\Congr$. More precisely, we would like to investigate
if, when this equivalence relation is a set of Tamari intervals (or is
closed by interval, or satisfies some other classical properties coming
from poset theory), one harvests a nice description of the Hilbert
series and of a combinatorial realization of $\Mag/_{\Congr}$.
\medbreak

A last research axis relies on the study on the $2$-magmatic operad
$\TwoMag$, that is, the free operad generated by two binary elements.
The analog of the associative operad in this context is the operad
$\TwoAs$~\cite{LR06} defined as the quotient of $\TwoMag$ by the
congruence saying that the two generators are associative. This operad
has a nice combinatorial realization in terms of alternating bicolored
Schröder trees. The question consists here in generalizing our main
results for the quotients of $\TwoMag$ and the generalizations of
$\TwoAs$ (that is, the definition of analogs of comb associative operads
and the study of their presentations).
\medbreak