### 2019-01-21 Response.tex
Response

% Author: Cyrille Chenavier, Christophe Cordero, Samuele Giraudo
% Creation: jan. 2019
% Modifications: jan. 2019

\documentclass[10pt,reqno]{amsart}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath,amsfonts,amssymb,amsthm,shuffle}
\usepackage[T1]{fontenc}
\usepackage[math]{anttor}

% Layout.
\usepackage[top=3.5cm,bottom=3.5cm,left=3.4cm,right=3.4cm]{geometry}

% Colors of hyperlinks.
\usepackage[dvipsnames]{xcolor}

\usepackage[hyperindex=true,frenchlinks=true,colorlinks=true,
citecolor=Col2,linkcolor=Col3,urlcolor=Col4,linktocpage,
pagebackref=true]{hyperref}

% Tikz.
\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}
\usetikzlibrary{decorations.pathmorphing}

% Misc.
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{wasysym}
\usepackage{stmaryrd}
\usepackage{cite}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{cancel}
\usepackage{xspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Line space.
\linespread{1.15}
\renewcommand{\arraystretch}{1.4}

% Vertical space for equations.
\setlength{\abovedisplayskip}{5pt}
\setlength{\belowdisplayskip}{5pt}

% Alphabetic footnote marks.
\renewcommand{\thefootnote}{\alph{footnote}}

% To allow cutting equations in several pages.
\allowdisplaybreaks

% Numbering of equations.
\numberwithin{equation}{subsection}

% Depth of the table of contents.
\setcounter{tocdepth}{2}

% Indentation in the table of contents.
\makeatletter
\def\l@section{\@tocline{1}{3pt}{1pc}{5pc}{}}
\def\l@subsection{\@tocline{2}{2pt}{2pc}{5pc}{}}
\makeatother

% Environments definitions.
\newtheorem{Theorem}{Theorem}[subsection]
\newtheorem{Proposition}[Theorem]{Proposition}
\newtheorem{Lemma}[Theorem]{Lemma}

% Better comparison symbols.
\renewcommand{\leq}{\leqslant}
\renewcommand{\geq}{\geqslant}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[Response to Referees]
    {Response to referee comments \\
    and \\
    changes report}
\keywords{}
\subjclass[2010]{}
\date{\today}
\author{Cyrille Chenavier \and Christophe Cordero \and Samuele Giraudo}
\address{\scriptsize Université Paris-Est, LIGM (UMR $8049$), CNRS,
    ENPC, ESIEE Paris, UPEM, F-$77454$, Marne-la-Vallée, France}
\email{cyrille.chenavier@u-pem.fr}
\email{christophe.cordero@u-pem.fr}
\email{samuele.giraudo@u-pem.fr}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\maketitle

Please find here the modifications made to the manuscript
``Quotients of the magmatic operad: lattice structures and convergent
rewrite systems'' submitted to Experimental Mathematics, following
the comments of the reviewer.
\bigbreak

\begin{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item `` {\it
The allusions in two places to "intervals in the Tamari order" are
rather strange. If you have any evidence or hint that this could be a
good condition, please tell the reader.
}''
\smallbreak

For the while, we have no general result connecting the quotients of
the magmatic operad and intervals of Tamari lattices but we suspect
that there is a link between the properties of the equated trees
with respect to the Tamari order and the description of the quotient
operad. This question was actually one of the starting points of this
work, and the research rather leads up to the presented results.
\medbreak

\item `` {\it
In my humble opinion, when doing so, one should cite carefully what 
is being used (language, package, etc), just like articles are cited.
}''
\smallbreak

We do not use a particular package. We added 
"All our computer programs are made from scratch in OCaml" in the 
introduction. 
\medbreak

\item `` {\it
The definition of $z_d$ page 20 needs to be checked (composition index)
}''
\smallbreak

The index was wrong. We changed "d/2" into "d-1/2".
\medbreak

\item `` {\it
About table of numbers page 23, do you see any interesting patterns ? 
possible formulas ? rationality of generating series ?
}''
\smallbreak

No ! Even with the package gfun on Mapple, we do not 
see any pattern, formulas, $\dots$
\medbreak

\item `` {\it
In section 4, the sentence "The reader can check" (repeated 4 times) 
is a bit terse. This is certainly feasible by hand. The reader would 
appreciate to be given hints of arguments for confluence (no critical 
pair ?) and termination. Or maybe there are available software that 
can swallow this kind of rewriting system and check that is is closed ?
}''
\smallbreak

We added the "using the Buchberger algorithm for operads" for the first 
occurrence of "The reader can check". It is easy to run the algorithm 
by hand on those examples.
\medbreak

\item `` {\it
Question: make a statement about the non-isomorphism (as set-operads 
or as k-linear-operads) between the 4 quotients indexed by compositions.
}''
\smallbreak

?
\medbreak

\item `` {\it
For the conjecture of section 4.3, up to which arity was it verified ?
}''
\smallbreak

Until arity 40 ! We added it in the article.
\medbreak

\end{itemize}

\end{document}